### 2018-02-18 doc/esm.Rmd
removed most of esm material, included ref to NEFA paper

---
output: fost::jlr_manuscript
---

```{r setup, include=FALSE}
devtools::load_all()
set_options()
knitr::opts_chunk$set(results = "asis", echo = FALSE)

# Captions
sfig <- captioner::captioner(prefix = 'Supplemental Figure')
cite_sf <- pryr::partial(sfig, display = 'cite')

tab <- captioner::captioner(prefix = 'Table')
cite_t <- pryr::partial(tab, display = 'cite')
stab <- captioner::captioner(prefix = 'Supplemental Table')
cite_st <- pryr::partial(stab, display = 'cite')

tab_qic <- stab('qic', 'Comparison of GEE model fitness for variable selection using quasi-likelihood information criteria.')
tab_tagfa <- stab('tagfa', 'Concentration (nmol/mL) and relative percent (mol%) values of triacylglycerol fatty acids in PROMISE participants at the baseline visit (2004-2006).')
tab_gee_unadj <- stab('gee_unadj', 'Raw estimates and confidence interval values for *time*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')
tab_gee_adj <- stab('gee_adj', 'Raw estimates and confidence interval values for *fully*-adjusted GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, waist circumference, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. Estimates represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with an asterisk (*) denoting a significant (p<0.05) association.')

fig_consort <- sfig('consort', 'CONSORT diagram of PROMISE participants over the 3 visits.')
fig_dag_is <- sfig('dag_is', 'Directed acyclic graphic output from the DAGitty online software for insulin sensitivity.')
fig_dag_bcf <- sfig('dag_bcf', 'Directed acyclic graphic output from the DAGitty online software for beta-cell function.')
fig_gee_adj_nowaist <- sfig('gee_adj_nowaist', 'Fully-adjusted (without waist size) GEE models of the association of the triacylglycerol fatty acids (mol% and nmol/mL) and total clinically-measured TAG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. X-axis values represent a percent difference in the outcome per SD increase in the fatty acid. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
```

# Supplemental Tables

```{r table_qic}
analyze_qic() %>% 
    table_qic(caption = tab_qic)
```

Given the number of possible combinations of outcome and predictor variables,
only ISI and ISSI-2 with total triacylglycerol fatty acids (nmol/mL) were used to
compare various GEE models and to select a final model.  Baseline age was used
as including both the original age and the time variable would result in
collinearity.  Column names: QIC is the quasi-likelihood information criteria
(smaller values, eg. larger negative values, indicate a better fit compared to
other models), Delta is the QIC minus the lowest QIC (models with delta <10 are 
considered equivalent).  Models were:

- M0: log(ISSI-2) or log(ISI) = total triacylglycerol fatty acids (nmol/mL) + years from baseline
- M1: M0 + fatty acid by time interaction
- M2: M0 + sex + ethnicity + baseline age
- M3: M2 + waist
- M4: M3 + ALT
- M5: M4 + physical activity (MET)
- M6: M5 + total NEFA
- M7: M6 + alcohol intake
- M8: M7 + family history of diabetes
- M9: M8 + smoking status

```{r table_tagfa}
table_tagfa(project_data, caption = tab_tagfa)
```

```{r table_gee}
table_gee_main(gee_results$unadj, caption = tab_gee_unadj)
table_gee_main(gee_results$adj, caption = tab_gee_adj)
```

# Supplemental Figures

![`r fig_consort`](../img/flowDiagramSample.png)

![`r fig_dag_is`](../img/dagitty-model-is.png)

![`r fig_dag_bcf`](../img/dagitty-model-bcf.png)

```{r fig_gee_adj_nowaist, fig.cap=fig_gee_adj_nowaist}
project_data %>% 
    analyze_gee(covars = covariates[which(covariates != 'Waist')]) %>% 
    mutate(p.value = ifelse(p.value <= 0.05, 0.04, p.value)) %>% 
    plot_gee_main() +
    ggplot2::theme(legend.position = "none")
```

```{r fig_pls_loadings}
p1 <- plot_pls_loadings(pls_results$homa2) +
    ggtitle("PLS model loadings for HOMA2-%S")
p2 <- plot_pls_loadings(pls_results$isi) +
    ggtitle("PLS model loadings for ISI")
p1 + p2 + plot_layout(ncol = 1)
```

```{r, eval=FALSE}
pls_results$homa2 %>% 
    plot_pls_scores() +
    theme_classic() +
    theme(strip.background = element_blank())
```