### 2018-04-26 doc/submission/reviewer-response.md
added first draft of responses to reviewers

---
output: fost::default_manuscript
---

<!-- add manuscript number to cover letter -->

<!--
The editorial staff will be in touch shortly to suggest revisions to your title
and abstract for you to consider. Please wait to hear from them before
submitting your revised manuscript. 
-->

# Editor

- All three reviewers agreed that the paper had the strength of longitudinal
followup and would be of interest to readers of the JLR but they also noted
that there have been earlier studies with similar outcomes and one reviewer
noted that some of those papers were by the present group but were not
mentioned here. The reviewers also thought that the discussion was too
speculative and the authors need therefore, to moderate their enthusiasm. My
additional comment is that although you quote one paper by Yki-Jarvinen and
colleagues. you did not discuss/reference her more recent papers linking
shorter chain saturated fatty acids in the liver to NAFLD/NASH. You should look
at those papers and add them to the discussion. I look forward to a modified
version.

    *We have made additions to the manuscript based on the editor's and reviewers'
    suggestions, including the discussion around short chain SFA and NAFLD by
    Yki-Jarvinen. We believe this has strengthened the paper and hope it
    satisfactorily addresses the editor and reviewers comments.*

# Reviewer 1

1. This manuscript reports interesting and potentially valuable associations
between fatty acid composition and metabolic endpoints in humans over time.

2. The associations are powerful and stand alone. The authors have no data
directly implicating de novo lipogenesis, so the last sentence in the abstract
about previous research should be removed or substantially modified.

    *We thank the reviewer for these comments. The suggestion is convincing, we 
    have removed the sentence from the abstract.*

3. There are statements in the Discussion that are overly speculative, dated,
and misleading. At the bottom of page 13 the phrase "may lead to greater DNL in
an attempt to control blood glucose" should be removed. On page 14, lines 8-9
the sentence "Higher insulin resistance may encourage greater DNL to handle the
higher blood glucose" should be removed. In the next sentence on the same page,
the authors discuss reference 47 by implying that weight loss improved fatty
acid profiles through effects on DNL, but that is facile and does not consider
the complex effects of improved insulin sensitivity on lipoprotein clearance
associated with induction of lipoprotein lipase, so this sentence needs to be
eliminated or modified.

    *These concerns and suggestions are appropriately raised. We have rephrased
    these discussions to be more cautious and less speculative.*

4. While the discussion emphasizes DNL, it does not allude to how DNL can
mechanistically lead to the metabolic endpoints used in this study. It would be
helpful to indicate that DNL is complicated, and in certain tissues is involved
in the promotion of chronic inflammation leading to decreased insulin
sensitivity and other adverse metabolic events (Wei X et al. Nature 2016;
539:294-298).

    *The discussion on DNL has been clarified in light of the reviewer's comments.*

5. On page 9, line 11 from the bottom, the text omits the actual r values for
the relationships between fatty acids and TG (should be in the 0.8 range for
many) and waist circumference (should be in the 0.3 range for many). More
omissions are made on lines 4 and 5 from the bottom when referring to the data
in Figure 3.

    *The r values have been included in the text {{confirm this is appropriate}}.*

# Reviewer 2

1.	It is a frustrating that this is seemingly the third in a series of papers
that is looking at patterns in lipid fractions and association with metabolic
outcomes. Thus, the TG, NEFA and phospholipid fractions all seem to be a 'least
publishable unit'. The authors must integrate the knowledge gained from the
previous analysis of NEFA and phospholipids and discuss some of the seemingly
contradictory findings from the previous studies. If the data were examined,
would the DNL story 'hold up'?

    *We thank the reviewer for these comments. We had many discussions over
    several months deciding how we can publish these results. Given the volume
    of results (figures, tables, numbers) and the complexity of the biology 
    involved in the present analysis, current journal article submissions are
    not suited for publishing a complete analysis and presentation of all the 
    results for all the lipid fractions into a single paper. We were not
    entirely satisfied with splitting the analyses up into multiple papers, 
    however it was the best course of action that we could see. The complete 
    and more integrated discussion and presentation of the results can be found
    in the first author's (LWJ) PhD thesis {{link}}. As to a combined analysis,
    we have presented these results at the European Diabetes Epidemiology Group
    Meeting in April, 2018 {{include DOI}}, and the DNL story remains and is in
    fact a strong contributor to fatty acid and diabetes pathogenesis associations.
    A short aside, these currently unpublished findings do not invalidate or 
    undermine the previous work. The combined analysis uses a powerful
    statistical technique (PLS), but it has a number of major limitations,
    not least of which is it can only handle cross-sectional data and can not
    adjust for potential confounders. The present analysis uses PLS in addition
    to the powerful longitudinal technique GEE. {{include a short 
    description expanding on these results here or in the manuscript?}}*

2.	The discussion of GEE and its limitations is well done. However, it seems
that the analysis suggests that total triglycerides are predictive of HOMA2-%S
and the other parameters. Do we need to subfractionate to get insight into risk
for type 2 diabetes?

    *This is excellent question. For a couple of reasons, no we don't need to 
    subfractionate TAG to understand risk. Firstly, we found that the four TGFA
    also highly positively correlated with clinical TAG. Secondly, clinical TAG
    had a much higher magnitude of association compared to any individual
    TGFA. Measuring clinical TAG is enough to assess risk. However, our results
    provide insight into the potential mechanisms that higher TAG influences risk
    for diabetes, which we show could be due to higher amounts of the four TGFA
    and that higher amounts of these four TGFA alone may strongly contribute to
    the higher clinical TAG.*

3.	Supplemental figures 2 and 3 are not discussed at all. They appear to be
interesting and informative in the development of the model. While the process
is described in the earlier NEFA paper, the authors should help the reader
understand what information is being provided without having to go back to a
previous paper.

    *We thank the reviewer for this comment. We agree, this information is useful
    and have added a brief discussion of it to the Methods section.*

# Reviewer 3

1. The authors noted that cluster of four TGFA that correlated negatively with
IS (14:0, 16:0, 14:1n-7, 16:1n-7) are products of DNL, and are therefore likely
reflecting a connection between DNL, CHO intake and diabetes risk. This
conclusion seems a bit overstated. Clearly, a link between this cluster and DNL
is possible, but the signature could be also be reflecting dietary intake,
hepatic steatosis and/or TGFA clearance by tissues such as muscle and adipose
tissue. The authors should include a more balanced discussion of potential
explanations/mechanisms beyond DNL.

    *We thank the reviewer for these comments. We have revised the discussion to 
    include alternative possibilities.*

2. Are there any additional analyses that can be offered to strengthen evidence
the TGFA composition is indeed a biomarker of DNL and/or CHO intake is this
specific cohort? For example, did the PROMISE study include food recall
records?

    *PROMISE did collect food frequency questionnaires, however this was done at
    the 3rd year visit while the fatty acids where measured at the baseline visit.
    So while we could technically use this data to help understand the data,
    we could also obtain inaccurate or biased results because of the time delay.
    We therefore decided to not analyze this data. {{run analysis and mention it
    here}}*

3. Also, the discussion should mention and consider that blood samples were
acquired after an overnight fast. Does fasting/fed state impact the correlation
between TGFA composition and DNL?

    *Excellent suggestion. We have included this in the discussion.*

Minor comments.

4. The authors should comment on how results reported here combine with
(consistencies/contradictions) those reported in their recent Diabetalogia paper
(the parallel NEFA analysis).

    *We have included a comment in the discussion on how our current analysis
    fits with our previous NEFA analysis.*

5. Authors underscore the strength of the statistical modeling used for the
analysis, including the use DAG modeling to identify potential confounders.
Readers might benefit from a brief comment on the basic principles and strengths
of the method.

    *We agree with this suggestion, but initially excluded it from the methods to
    limit word count. We have included a section in the methods that briefly 
    explains this method.*