### 2018-06-05 doc/manuscript.Rmd
minor edits and typo fixes in manuscript after final proofreading

---
bibliography: 
    - /home/luke/Documents/bibliography/bib/bib-db/tagDiabetes.bib
    - /home/luke/Documents/bibliography/bib/bib-db/myrefs.bib
    - /home/luke/Documents/bibliography/bib/bib-db/methods.bib
    - /home/luke/Documents/site/cv/_includes/work.bib
csl: /home/luke/Documents/bibliography/styles/journal-of-lipid-research.csl
output: fost::jlr_manuscript
---

# Title page

Title (110 / 110 characters): 
Clusters of fatty acids in the serum triacylglyceride fraction associate with the
disorders of type 2 diabetes

Authors: Luke W. Johnston, PhD (1); Zhen Liu, PhD (1); Ravi Retnakaran, MD
(2,3); Bernard Zinman, MD (2,3); Adria Giacca, PhD (4); Stewart B. Harris, MD (5);
Richard P. Bazinet, PhD (1); and Anthony J. Hanley, PhD (1,2,6)

Institutions:

(1) Department of Nutritional Sciences, University of Toronto, Toronto, Ontario,
Canada.
(2) Leadership Sinai Centre for Diabetes, Division of Endocrinology, University
of Toronto, Toronto, ON, Canada.
(3) Lunenfeld Tanenbaum Research Institute, Mount Sinai Hospital, Toronto,
Ontario, Canada.
(4) Department of Physiology, University of Toronto, Toronto, Ontario, Canada.
(5) Centre for Studies in Family Medicine, University of Western Ontario,
London, Ontario, Canada.
(6) Dalla Lana School of Public Health, University of Toronto, Toronto, Ontario,
Canada.

Corresponding author contact information:

- Name: Anthony J. Hanley
- Current address:  
    Department of Nutritional Sciences  
    Faculty of Medicine  
    University of Toronto  
    FitzGerald Building, 150 College Street, Room 341  
    Toronto, ON, Canada, M5S 3E2
- Phone number: 416.978.3616
- Email: anthony.hanley@utoronto.ca

Running title (54 / 60 characters):
Triacylglyceride fatty acids and diabetes pathogenesis

# Abbreviations:

- ALT: Alanine aminotransferase
- CV: Coefficient of variation
- DNL: De novo lipogenesis
- DAG: Directed acyclic graph
- GEE: Generalized estimating equations
- HOMA2-%S: Homeostatic model of assessment 2 - percent sensitivity
- IGI/IR: Insulinogenic index over homeostatic model of assessment for insulin
resistance
- IS: Insulin sensitivity
- ISI: Insulin sensitivity index
- ISSI-2: Insulin Secretion-Sensitivity Index-2
- MAQ: Modified activity questionnaire
- MET: Metabolic equivalent of task
- NEFA: Non-esterified fatty acids
- OGTT: Oral glucose tolerance test
- PROMISE: Prospective Metabolism and Islet Cell Evaluation cohort
- PLS: Partial least squares
- SD: Standard deviation
- TGFA: Triacylglycerol fatty acids
- WC: Waist circumference

# Abstract

**Aims:**
Our aim was to examine longitudinal associations of triacylglyceride fatty acid
(TGFA) composition with insulin sensitivity (IS) and beta-cell function.

**Methods**:
Adults at-risk for T2D (n=477) had glucose and insulin measured from a glucose
challenge at 3 time points over 6 years. The outcome variables Matsuda index
(ISI), HOMA2-%S, Insulinogenic Index over HOMA-IR (IGI/IR), and Insulin
Secretion-Sensitivity Index-2 (ISSI-2) were computed from the glucose challenge.
Gas chromatography quantified TGFA composition from the baseline. We used
adjusted generalized estimating equations (GEE) models and partial least squares
(PLS) regression for the analysis.

**Results**: 
In adjusted GEE models, four TGFA (14:0, 16:0, 14:1n-7, 16:1n-7 as mol%) had
strong negative associations with IS while others (e.g. 18:1n-7, 18:1n-9,
20:2n-6, 20:5n-3) had strong positive associations. Few associations were seen
for beta-cell function, except for 16:0, 18:1n-7, and 20:2n-6. PLS analysis indicated four
TGFA (14:0, 16:0, 14:1n-7, 16:1n-7) that clustered together and strongly related
with lower IS. These four TGFA also correlated highly (r>0.4) with clinically
measured TG.

**Conclusions**:
We found that higher proportions of a cluster of four TGFA strongly related with
lower IS as well as hypertriglyceridemia, suggesting only a few fatty acids within
the TGFA composition may primarily explain lipids' role in glucose dysregulation.

# Introduction

Hypertriglyceridemia is a well described metabolic disorder resulting in
negative health outcomes [@Chehade2013a]. It is a risk factor for cardiovascular
disease [@DAgostino2004a; @Verges2015a] and is associated with other metabolic
disorders such as non-alcoholic fatty liver disease [@Kawano2013a], the
metabolic syndrome [@Alberti2009], and abdominal obesity [@Lemieux2000a].
Circulating triacylglceride (TG) concentration is commonly measured during
routine clinical assessment using enzymatic methods. However, clinically
measured TG is limited as it represents the full fatty acid spectrum within the
TG fraction as a summary measure. There is increasing appreciation for the
importance of specific fatty acid composition profiles in different plasma
fractions on various health outcomes [@Forouhi2014a; @Ma2015a; @Johnston2016a;
@Johnston2017b], however there are relatively few studies that have explored the
impact of the fatty acid composition in the TG fraction [@Rhee2011a;
@Lankinen2015a].

The interaction between TG and insulin sensitivity is complex and involves
components of a feedback system [@Verges2015a]. Greater resistance to insulin in
both the liver and muscle may result in greater production of TG and secretion
of lipoproteins that transport TG [@Yu2012a]. Likewise, greater TG may
contribute to metabolic dysfunction and lipotoxicity in various tissues,
affecting insulin sensitivity, and thus continue the cycle [@Verges2015a]. Given
the complexity and temporal nature of the relationship, long term studies with
multiple data collection time points are paramount to better understanding the
underlying biology and subsequent risk.

While several studies have documented prospective associations of
hypertriglyceridemia with incident type 2 diabetes (T2D) [@Chien2008a;
@DAgostino2004a; @Schulze2009a], only a limited number of longitudinal studies
[@Rhee2011a; @Lankinen2015a] have examined the relationship between TG and its
composition with the pathophysiological factors underlying T2D,
particularly beta-cell function. Our objective was to examine the longitudinal
role of the specific composition of the serum TG fraction on OGTT-derived
measures of insulin sensitivity and beta-cell function compared to clinically
measured TG in a Canadian population at risk for T2D.

# Materials and Methods

```{r setup, message=FALSE, echo=FALSE}
devtools::load_all('.')
library(ggplot2)
library(patchwork)
library(dplyr)
knitr::opts_chunk$set(echo = FALSE)
ds0yr <- project_data %>% 
    filter(VN == 0)

## Include captions below using `captioner` package
fig <- captioner::captioner(prefix = 'Figure')
cite_f <- pryr::partial(fig, display = 'cite')
sfig <- captioner::captioner(prefix = 'Supplemental Figure S')
cite_sf <- pryr::partial(sfig, display = 'cite')

tab <- captioner::captioner(prefix = 'Table')
cite_t <- pryr::partial(tab, display = 'cite')
stab <- captioner::captioner(prefix = 'Supplemental Table S')
cite_st <- pryr::partial(stab, display = 'cite')
# usage: cite_st('basicChar')

tab_basic <- tab('basic', 'Basic characteristics of PROMISE participants at each of the 3 clinic visits.')

fig_tagfa <- fig('tagfa', 'Distribution of the composition of TGFA in the baseline visit of PROMISE participants (2004-2006). Boxplots represent the median and interquartile range of the FA values.')
fig_heatmap <- fig('heatmap', 'Pearson correlation heatmap of TGFA (nmol/mL) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_mol <- fig('heatmap_mol', 'Pearson correlation heatmap of TGFA (mol%) with continuous basic and metabolic characteristics of PROMISE participants from the baseline visit (2004-2006). Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_heatmap_tagfa <- fig('heatmap_tagfa', 'Pearson correlation heatmap of TGFA in the PROMISE participants from the baseline visit (2004-2006). The correlations of FA grouped using heirarchical cluster analysis; FA along the x and y axis are ordered according to this analysis. Darker orange represents a positive correlation; darker blue represents a negative correlation.')
fig_gee_unadj <- fig('gee_unadj', 'Time-adjusted GEE models of the association of the TGFA (mol% and nmol/mL) and total clinically-measured TG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. X-axis values represent a percent difference in the outcome per SD increase in the FA. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_gee_adj <- fig('gee_adj', 'Fully-adjusted GEE models of the association of the TGFA (mol% and nmol/mL) and total clinically-measured TG with insulin sensitivity and beta-cell function outcomes using the 6 year longitudinal data from the PROMISE cohort. Variables controlled for were follow-up time, WC, baseline age, ethnicity, sex, ALT, physical activity, and total NEFA. X-axis values represent a percent difference in the outcome per SD increase in the FA. P-values were adjusted for the BH false discovery rate, with the largest dot representing a significant (p<0.05) association.')
fig_pls <- fig('pls', 'Partial least squares (PLS) models showing the clustering of TGFA on insulin sensitivity and beta-cell function measures. The percent explained variance of each component is shown in brackets on each axis. The solid line represents an explained variance of 100% while the dashed line represents an explained variance of 50%. FA between these lines represent variables that strongly explain the underlying structure of the data.')
```

```{r analyze_others}
# Correlation between predicted and observed
pls_homa2_t <- analyze_pls(y = 'lHOMA2_S')
pls_isi_t <- analyze_pls(y = 'lISI')
pls_igi_t <- analyze_pls(y = 'lIGIIR')
pls_issi2_t <- analyze_pls(y = 'lISSI2')

pls_gee <- project_data %>%
    analyze_gee(
        x = c('Comp1', 'Comp2'),
        data_prep = 'pls',
        pls_results = pls_results$isi
        ) %>% 
    dplyr::select(Yterms, Xterms, estimate, p.value) %>% 
    dplyr::filter(Xterms == 'Comp1') %>% 
    dplyr::mutate(p.value = aide::format_pval(p.value),
                  estimate = aide::format_round(estimate))
```

```{r inline}
tagfa <- calc_tagfa_percent()
outcome_chg <- calc_outcome_changes()
female <- calc_discr_npct(ds0yr$Sex, 'Female')
euro <- calc_discr_npct(ds0yr$Ethnicity, 'European')
fhd <- calc_discr_npct(ds0yr$FamHistDiab, 'Yes')
gmest0 <- calc_gee_estci(gee_results$unadj)
gmest1 <- calc_gee_estci(gee_results$adj)
fa_cor <- calc_cor(cor_results$conc)
fa_cor_mol <- calc_cor(cor_results$pct)
cor_pls_homa2 <- calc_pred_corr(pls_homa2_t, pls_homa2_t$test_data)
cor_pls_isi <- calc_pred_corr(pls_isi_t, pls_isi_t$test_data)
cor_pls_igi <- calc_pred_corr(pls_igi_t, pls_igi_t$test_data)
cor_pls_issi2 <- calc_pred_corr(pls_issi2_t, pls_issi2_t$test_data)
```

<!-- Use an external source via DOI for reference to these methods? -->

Recruitment for the baseline visit of the Prospective Metabolism and Islet Cell Evaluation (PROMISE) 
cohort took place between 2004-2006 in London and Toronto, 
Canada. Individuals were selected to participate if they met the eligibility 
criteria of having one or more risk factors for T2D, 
including obesity, hypertension, family history of diabetes, and/or a history of
gestational diabetes or birth of a macrosomic infant. A total of 736 individuals
attended the baseline visit. Subsequent examinations occurred every three years,
with data from three examination visits available for the present analysis
(2004-2006, 2007-2009, and 2010-2013). The current study
used data on participants who did not have T2D at baseline, who
returned for one or more of the follow-up examinations, and who had samples
available for fatty acid measurements (n=477; see the CONSORT diagram in 
`r cite_sf('consort')`). Metabolic characterization, anthropometric
measurements, and questionnaires on lifestyle and sociodemographics were
administered at each examination visit. Research ethics approval was obtained from
Mount Sinai Hospital and the University of Western Ontario, and all participants
provided written informed consent. Data collection methods were standardized
across the 2 centres and research nurses were centrally trained.

## Metabolic characterization

After 8-12 hours of overnight fasting, participants completed a 75g oral glucose tolerance test (OGTT) at 
each examination visit, with blood samples taken at fasting, 30 min, and 2 hr 
post-glucose load. Samples were subsequently processed and frozen at -70°C.
Alanine aminotransferase (ALT) was measured using standard laboratory 
procedures. Cholesterol, HDL, and clinically-measured TG 
were measured using Roche Modular's enzymatic colorimetric tests (Mississauga, 
ON). Both insulin and glucose were
measured from OGTT blood samples at fasting, 30 minute,
and 2 hour time points. Specific insulin was measured with the Elecsys 1010 
(Roche Diagnostics, Basel, Switzerland) immunoassay analyzer and 
electrochemiluminescence immunoassay, which shows 0.05% cross-reactivity to 
intact human pro-insulin and the Des 31,32 circulating split form (Linco Res. 
Inc) and has a coefficient of variation (CV) of 9.3%. Glucose was determined 
using an enzymatic hexokinase method (Roche Modular, Roche Diagnostics) with a 
detection range of 0.11 to 41.6 mmol/L and an inter-assay CV of <1.1% and an
intra-assay CV of < 1.9%.  All assays were performed at the Banting and Best
Diabetes Centre Core Lab at Mt Sinai Hospital.

TGFA composition was quantified using stored fasting serum samples from the 
baseline visit, which had been frozen at -70°C for 4-6 years and had not been 
exposed to any freeze-thaw cycles. Serum fatty acids have been documented to be 
stable at these temperatures for up to 10 years [@Matthan2010a]. A known amount 
of triheptadecanoin (17:0; Nu-Chek Prep, Inc Elysian, MN, USA) was added as an internal
standard prior to extracting total lipids according to the method of Folch et al
[@Folch1957a]. Each serum lipid fraction (non-esterified fatty acids (NEFA), 
cholesteryl ester, phospholipid, and TG) was isolated using thin layer 
chromatography. Fatty acid methyl esters were separated and quantified using a 
Varian-430 gas chromatograph (Varian, Lake Forest, CA, USA) equipped with a 
Varian Factor Four capillary column and a flame ionization detector. Fatty acid 
concentrations (nmol/ml) were calculated by proportional comparison of gas
chromatography peak areas to that of the internal standards [@Nishi2014a]. There
were 22 fatty acids measured in the TGFA fraction. 
Findings for other lipid fractions in this cohort are reported separately (see
[@Johnston2016a] for the phospholipid and cholesteryl ester fraction and
[@Johnston2017b] for the NEFA fraction analysis).

## Anthropometrics and sociodemographics

Height, weight, and waist circumference (WC) were measured at all clinic 
examinations using standard procedures.  WC was measured at the natural waist, 
defined as the narrowest part of the torso between the umbilicus and the xiphoid
process. BMI was calculated by dividing weight (kg) by height (m) squared.
Questionnaires administered at each examination determined sociodemographics.
A version of the Modifiable Activity Questionnaire (MAQ) [@Kriska1990]
determined estimated physical activity. The MAQ collects information on leisure
and occupational activity, including intensity, frequency, and duration, over
the past year. Each reported activity from the MAQ was weighted by its
metabolic intensity allowing for the estimation of metabolic equivalents of
tasks (MET) hours per week [@Kriska1990].

## Variable calculation and statistical analysis

Insulin sensitivity and beta-cell function indices were computed using the OGTT 
glucose and insulin data. Insulin sensitivity was assessed using the Insulin
Sensitivity Index (ISI) [@Matsuda1999] and HOMA2-%S [@Levy1998a] using the HOMA2
Calculator. HOMA largely reflects hepatic insulin sensitivity, while ISI
reflects whole-body insulin sensitivity [@AbdulGhani2007]. Beta-cell function
was assessed using the Insulinogenic Index [@Wareham1995] over HOMA-IR [@Matthews1985] (IGI/IR)
and the Insulin Secretion-Sensitivity Index-2 (ISSI-2) [@Retnakaran2009]. IGI/IR
is a measure of the early phase of insulin secretion while ISSI-2 is analogous to
the disposition index (but is calculated using OGTT values). Each index has been validated
against gold standard measures [@Matthews1985; @Hermans1999a; @Matsuda1999; @Retnakaran2009].

The primary outcome variables for this analysis were HOMA2-%S, ISI, IGI/IR, and
ISSI-2, which were log-transformed for the statistical modeling. The primary
predictor variables for this analysis were 22 individual TGFA included as either mole percent
(mol%) of the total fraction or as a concentration (nmol/mL). Clinically-measured
TG was also included as a primary predictor to allow us to test the hypothesis
that specific TGFA better predicted outcomes compared to clinical TG. 
Pearson correlation coefficients were computed to assess the relationships of
individual TGFA with other continuous variables. Within TGFA composition correlations were also
computed and subsequently analyzed using hierarchical clustering.

Generalized estimating equation (GEE) models [@Zeger1986a] were used in the 
primary analysis to determine the longitudinal associations between the outcome 
variables and the predictor variables. The predictor variables
were scaled (mean-centered and standardized). Given the longitudinal 
design, an auto-regressive of order 1 working correlation matrix was 
specified in the GEE model. Covariates to adjust for were selected based on the
previous literature, from directed acyclic graph (DAG) [@Greenland1999a]
recommendations, and from quasi-likelihood information criteria. 
DAGs are used to identify the minimum adjustment necessary for a model by
using the causal pathways to algorithmically identify potential confounding and
colliding variables (see [@Greenland1999a] for more detail about using
DAGs). The DAG structures to understand potential confounding, shown in 
`r cite_sf("dag_is")` and `r cite_sf("dag_bcf")`, were processed by the DAGitty
software [@Textor2011a;@Shrier2008a] to generate the recommended adjustments.
These DAG structures were developed based on hypothesized causal pathways
between each variable, which were then input into the DAGitty software. The
output from DAGitty was used, in conjunction with the other methods, to help inform the final model.

The final GEE model (M6; seen in `r cite_st('qic')`) was adjusted for years
since baseline, WC, baseline age, ethnicity,
sex, ALT, MET, and total NEFA. The variables TGFA, total NEFA, sex, ethnicity, and
baseline age were classified as *time-independent* (held constant) as they were
measured only at the baseline visit or do not change throughout the study, while
the outcome variables and remaining covariates were set as *time-dependent*.
After transformations, the GEE estimates are interpreted as an expected percent 
difference in the outcome variable for every standard deviation (SD) increase in
the predictor variable given the covariates are held constant (including time). 
We also tested for an interaction with sex, ethnicity, or time by the predictor 
term for each outcome variable.

While GEE accounts for the longitudinal design of the data, this approach is 
limited in that it cannot analyze the inherent multivariate nature of the 
composition of the TGFA fraction. Therefore, to confirm the GEE results in a multivariate 
environment (i.e. all TGFA analyzed collectively), partial least squares regression (PLS) was
used to identify the patterns of TGFA composition against insulin sensitivity
and beta-cell function as outcome variables. 
Briefly, PLS is a technique that extracts latent
structures (clusters) underlying a set of predictor variables conditional on a
response variable(s) (i.e. the outcome variables). How accurately the clusters
within the TGFA composition predict metabolic function is determined
by using cross-validation on the PLS models.

A more detailed explanation of these statistical techniques and on the analysis
process can be found in the supplemental methods for our paper in the NEFA
fraction [@Johnston2017b].
All analyses were performed using R `r getRversion()` [@Rbase], along with the R
packages geepack `r packageVersion('geepack')` for GEE [@Hoejsgaard2006a] and pls
`r packageVersion('pls')` for PLS. The R code and extra analyses for
this manuscript is available at https://doi.org/10.6084/m9.figshare.5143438. Results were considered
statistically significant at p<0.05, after adjusting for multiple testing using
the Benjamini-Hochberg False Discovery Rate [@Benjamini1995a]. STROBE was used
as a guideline for reporting [@Vandenbroucke2007b].

# Results

## Basic characteristics of the PROMISE cohort

`r cite_t('basic')` shows basic characteristics of the PROMISE cohort. The mean 
follow-up time was `r calc_followup_time()` years, where `r calc_pct_full_visits()`%
of participants attended all three visits. There were
`r female$npct` females and `r euro$npct` who reported European-ancestry, with a mean
age in years of `r aide::ave_sd(ds0yr$Age)` and a mean BMI
of `r aide::ave_sd(ds0yr$BMI)` kg/m^2^. As expected from the study's eligibility
criteria, the majority of participants, n=`r fhd$npct`, had a family history of
diabetes. Between the baseline visit and the 6-year visit in this sample,
insulin sensitivity and beta-cell function measures had a significant median decline of
between `r outcome_chg$chg` (p`r outcome_chg$p` from GEE; n=`r outcome_chg$n`).

`r cite_f('tagfa')` shows the composition of each FA in the TG fraction (see 
`r cite_st('tagfa')` for the raw values). 
Three TGFA contributed `r aide::format_round(sum(tagfa$pct))`% to the total 
TG concentration: `r tagfa$c[1]`; `r tagfa$c[2]`; and, `r tagfa$c[3]`.
`r cite_f('heatmap')` shows a heatmap of the correlation of individual TGFA as
concentrations with the outcome variables and several basic characteristics.
As expected, nearly all TGFA had very strong positive correlations (r=
`r fa_cor$rng$TG[2]`) with clinically-measured TG and moderate positive 
correlations with WC (r=`r fa_cor$rng$WC[2]`). There were also moderate
negative correlations with HDL (r=`r fa_cor$rng$HDL[1]`). For the outcome variables,
the correlations for the insulin sensitivity measures were generally higher 
(HOMA2-%S: r=`r fa_cor$rng[['HOMA2-%S']][1]`, ISI: r=`r fa_cor$rng$ISI[1]`) than 
for the beta-cell function measures (all r<0.30). For correlations
of individual TGFA using mol% with the basic participant characteristics, as shown in 
`r cite_f('heatmap_mol')`, differences in correlations between fatty acids were
most evident for 14:0, 14:1n-7, 16:0, and 16:1n-7 that had a moderate positive
correlation with clinical TG (r=`r fa_cor_mol$rng$TG[2]`) while all other fatty
acids had a negative association (r=`r fa_cor_mol$rng$TG[1]`). In particular,
those fatty acids with the negative associations with clinical TG were all the
very long chain polyunsaturated fatty acids (e.g. 20:4n-6, 20:5n-3).
As seen in `r cite_f('heatmap_tagfa')`, four fatty acids (14:0,
16:0, 14:1n-7, and 16:1n-7) clustered together, each highly positively correlated
with each other and negatively correlated with all other fatty acids.

## Generalized estimating equation models

Results from the unadjusted GEE model are shown in `r cite_f('gee_unadj')` and 
for the adjusted GEE model in `r cite_f('gee_adj')`. The majority of
associations with beta-cell function measures were attenuated after full model
adjustment, while nearly all associations with insulin sensitivity remained
significant for both mol% and nmol/mL results. Subsequent analysis revealed that
the attenuation with beta-cell function was due primarily to adjustment for
WC.

In analyses using concentration values, nearly all TGFA had a strong
negative association on HOMA2-%S and ISI (estimates of percent difference ranging from
`r gmest1$outcome_b$rng[3]` and `r gmest1$outcome_b$rng[7]`, respectively), and a
few had strong negative associations with IGI/IR and ISSI-2 (estimates ranging 
from `r gmest1$outcome_b$rng[11]` and `r gmest1$outcome_b$rng[15]`, respectively).
In analyses using TGFA mol% values, four TGFA (14:0, 16:0, 14:1n-7, and 16:1n-7) had negative
associations with HOMA2-%S and ISI (estimates between `r gmest1$outcome_b$rng[1]` and 
`r gmest1$outcome_b$rng[5]`, respectively, lower insulin sensitivity for every SD
increase in the TGFA), while several more TGFA had positive associations with
HOMA2-%S and ISI (20:0, 18:1n-9, 20:1n-9, 22:1n-9, 18:2n-6, 20:2n-6, 20:4n-6,
and 22:5n-3) estimating between `r gmest1$outcome_b$rng[2]` and 
`r gmest1$outcome_b$rng[6]`%, respectively, higher insulin sensitivity for every
SD increase in the TGFA. One TGFA, 20:2n-6, had a very strong positive
association with the insulin sensitivity measures, with a `r gmest1$b_ci$est[1]`
to `r gmest1$b_ci$est[37]`% higher insulin sensitivity for every SD increase. 
Both clinically-measured TG and total TGFA concentration had very strong
negative associations with all outcome variables.

While there were a few
significant interactions by time in unadjusted models, after inclusion of
covariates in the model, these interactions were attenuated (data not
shown). There were no significant interactions by sex or ethnicity for any of
the TGFA (data not shown). Results of the sensitivity analyses identifying
WC as the covariate that attenuated the beta-cell function
associations from the unadjusted model are shown in `r cite_sf("gee_adj_nowaist")`. 
A tabular presentation of the GEE results is shown in `r cite_st("gee_unadj")`
for unadjusted models and `r cite_st("gee_adj")` for adjusted models.

## Clustering of TGFA by metabolic measures

The PLS analysis corroborated the findings from the GEE models. 
The PLS results conditioned on insulin sensitivity as the outcome showed a
clustering of the fatty acids 14:0, 14:1n-7, 16:0, and 16:1n-7 as mol%
(`r cite_f('pls')`). These TGFA loaded strongly and negatively on HOMA2-%S and ISI
in the first component, suggesting this cluster of TGFA tracks
together with lower insulin sensitivity. The TGFA 20:2n-6, 20:5n-3, 22:5n-3,
and 22:6n-3 loaded positively on both insulin sensitivity measures. No other
TGFA loaded strongly. In the second component, 18:1n-9 and 18:1n-7 loaded positively 
but not strongly while 20:5n-3 and 22:6n-3 loaded strongly and negatively with both
HOMA2-%S and ISI; however, this component only explained <10% of the variance. 
The PLS model for insulin sensitivity had good predictive
ability, with a high correlation between the predicted outcome values against
the observed values (HOMA2-%S: `r cor_pls_homa2$r_p`; ISI: `r cor_pls_isi$r_p`).

The beta-cell function PLS results showed a similar clustering of fatty acids,
however there was a lower correlation (though significant at 
p`r cor_pls_igi$p`) between the predicted values and the observed values 
(r=`r cor_pls_igi$r` to `r cor_pls_issi2$r`), suggesting that TGFA composition
poorly predicts beta-cell function.
Given the low predictability, only the insulin sensitivity measures are
presented. We used the extracted PLS scores as the predictor variable in the GEE
models and found negative associations of the first component on all outcome
variables, with the strongest association being with the insulin sensitivity
variables (estimate of `r pls_gee$estimate[3]`, all p`r pls_gee$p.value[1]`; using PLS
scores constrained by ISI). See `r cite_sf("pls_loadings")` for a plot of the
loadings of each TGFA on the two components.

# Discussion

In the present study, we found that in a Canadian cohort at risk for T2D,
several specific TGFA and groups of TGFA were strongly associated
with insulin sensitivity and moderately associated with beta-cell function. In 
particular, the TGFA myristic acid (14:0), 7-tetradecenoic acid 
(14:1n-7), palmitic acid (16:0), and palmitoleic acid (16:1n-7) all strongly and
negatively associated with lower insulin sensitivity. While most TGFA were not
associated with beta-cell function, three fatty acids, palmitic acid (16:0),
*cis*-vaccenic acid (18:1n-7), and eicosadienoic acid (20:2n-6) were associated negatively and positively,
respectively, with measures of beta-cell function. Using PLS, we also found that four
TGFA (14:0, 14:1n-7, 16:0, 16:1n-7) clustered together, and that this cluster
strongly predicted lower insulin sensitivity.

To our knowledge, no longitudinal study to date has examined the role of the
composition of the TGFA fraction on detailed OGTT-derived metabolic measures. 
Two large prospective studies have been published that similarly examined TGFA 
composition and T2D outcomes. Rhee *et al* presented a nested case-control
analysis (n=189 cases and n=189 controls) within the Framingham offspring cohort
[@Rhee2011a], which found that subjects with a TGFA composition characterized
by a lower carbon chain and fewer double bonds (e.g. 14:0, 16:0) had a higher
risk for T2D after 12-years while those with a profile characterized by
higher carbon chain and more double bond TGFA had a lower risk for T2D.
A similar pattern of TGFA was also associated with HOMA-IR cross-sectionally
at the baseline visit. In addition, Lankinen *et al* reported on
a prospective cohort of males in Finland [@Lankinen2015a], for which 
TGFA data were available for 831 participants after 6-years of follow-up. In
their cohort, OGTT data were only available at the 6-year visit.
In the cross-sectional analysis they found that most saturated fatty acids had negative 
associations with insulin sensitivity and beta-cell function while
linoleic acid (18:2n-6), docosapentaenoic acid (22:5n-3), eicosapentaenoic acid
(20:5n-3), and arachidonic acid (20:4n-6) had positive associations with insulin
sensitivity. The magnitude of the associations were larger for the insulin
sensitivity results compared to the beta-cell function results, similar to what
we observed. Our study extends these findings by using multiple measurements
of metabolic function and as well as multivariate statistical approaches that
allowed us to identify clusters of TGFA. In another much smaller study (n=16) of
mostly females [@Kotronen2009a], the authors reported a
positive correlation between total esterified (of which TG make up the
majority) 16:0, 16:1n-7, and 18:1n-9 with HOMA-IR, findings which were largely 
similar to the present analysis.

There are a few possible explanations for these findings. Circulating 
TGFA derive from three sources: adipose lipolysis, dietary fat, and de novo
lipogenesis (DNL). Dietary carbohydrates and fat can influence DNL activity
[@Aarsland1998a;@Harding2015a;@Luukkonen2018a]. Determining the specific source
of TGFA is extremely difficult to ascertain outside of highly controlled
experimental settings. Many previous studies that have examined DNL as the source
have used markers of estimated DNL, such as the ratio between 18:2n-6 to 16:0 or
16:1n-7 to 16:0 [@Hodson2008a; @Lankinen2015a; @Kroger2011a]. However, there are
limitations to using these ratios as the fatty acids used in their calculation
can also be obtained from the diet in addition to being created through DNL
[@Hodson2008a]. An experimental feeding trial (n=24) was conducted to identify the fatty acids
that most accurately reflected DNL as potential biomarkers [@Lee2015a]. The
study found that palmitoleic acid (16:1n-7), directly measured DNL using
isotopes, and liver fat were all highly correlated with each other (r>0.50),
suggesting that 16:1n-7 may be a good biomarker for hepatic DNL. In another
small (n=14) feeding trial, meal type (high fat vs low fat) was tested to
determine its effect on DNL and TGFA composition [@Wilke2009a]. The authors
reported that 14:0, 16:0, 16:1, and 18:2 were higher in the low fat (high
carbohydrate) group. In another recent overfeeding trial, 1000 kcal of saturated
fat, unsaturated fat, or carbohydrates over 3 weeks was given to 38 overweight
individuals [@Luukkonen2018a] to test changes in liver fat and hepatic DNL. At the
end of the study, the carbohydrate group had higher DNL activity as well as an increase
in liver fat, though the saturated fat group had the highest increase in liver
fat. This link between carbohydrate intake and DNL activity has been well documented
[@Hodson2008a; @Kawano2013a; @Harding2015a; @Hudgins2000a; @Parks1999a; @Chehade2013a; @Verges2015a].

In our findings, the four fatty acids were highly positively correlated
amongst each other and negatively or neutrally with all other TGFA, in addition to
clustering together on their negative association with insulin sensitivity. This
may suggest that greater DNL activity is the source of these TGFA. Several
studies have shown a link between higher estimated DNL activity and an increased risk for
metabolic dysfunction [@Ma2015a; @Zong2013a; @Lankinen2015a; @Kroger2011a]. How
DNL may influence metabolic dysfunction is not well understood. Possible
reasons may be that higher DNL produces more of certain fatty acids or that
higher DNL increases circulating TG, which itself is well documented to
contribute to metabolic dysfunction and which we found in our study as a high
positive correlation between the four TGFA and clinical TG.

Regardless of the exact source of these fatty acids, our results, in addition to
the available scientific evidence, emphasize the importance of the fatty acid
composition on metabolic health, as individual fatty acids can have specific
physiological functions. For instance, a higher concentration of
circulating 14 and 16 carbon fatty acids may expose tissues to greater
lipotoxicity, for instance from palmitic acid (16:0), which is well-known to
have harmful effects on tissues [@Riserus2008a; @Iggman2010a]. Our study extends
these findings by showing that TGFA with 14 to 16 carbons clustered together and
this pattern strongly associated with lower insulin sensitivity. While some of these fatty
acids also had a significant association with beta-cell function, the magnitude
of associations were more modest compared to those for insulin sensitivity.

The direction of association between TGFA and insulin sensitivity is unclear
from previous cross-sectional studies due to the physiological feedback
mechanisms involved. For example, while higher TGFA may promote
muscle insulin resistance, the reverse may also be true [@Flannery2012a].
As we found no interaction by time of TGFA on insulin sensitivity, this study
cannot determine the exact role of the feedback mechanism. However, combining
the lack of a time interaction and the consistent negative association
in models without the time interaction, these results at least suggest that the feedback
mechanism may not be strongly influential and that TGFA may predict insulin
sensitivity at least over a six year period. Given the complex biological mechanisms and
feedback loops involved, disentangling whether insulin sensitivity influences
TGFA more strongly than TGFA influencing insulin sensitivity will require more
complex research designs and analyses.

Given the close biological relationship between circulating NEFA and TG, NEFA 
may act as a confounding factor and was thus adjusted for. In our published
analysis of the NEFA fraction [@Johnston2017b], we found that higher total NEFA,
but not the specific composition, associated with lower beta-cell function. This
is in contrast to the TGFA findings that the specific composition does
differentially associate with insulin sensitivity and beta-cell function,
adjusting for total NEFA. There was no difference in results in models that did
not include NEFA as a confounder (data not shown). This difference in results
between NEFA and TGFA suggests that TGFA may independently and strongly influence the 
pathophysiology of T2D, when compared to other lipid fraction compositions,
including the phospholipid and cholesteryl ester fractions [@Johnston2016a]. This
may be due to TG being biologically destined for uptake by non-hepatic tissue as
they are found mainly in VLDL, at least during fasting. This is in contrast to
NEFA that are mostly taken up by the liver and used in TG production [@Nielsen2012a].

Our study has potential limitations that need to be considered when interpreting
the results. Firstly, this is an observational cohort and as such there may be
some residual confounding we were not able to control for or were unaware of. 
However, we have taken extensive, empirically based precautions in identifying
potential confounders and mediators through the use of the DAG modeling, relying
on previous literature, and through information criteria model fit comparison methods.
Only fasting TGFA were quantified, and only at the baseline visit. TGFA 
composition can fluctuate substantially throughout the day, so in order to 
control for this, PROMISE participants came for the clinic visit in the morning
and fasted. There is some evidence to suggest that fasting TG is better able
to discriminate diabetes cases compared to a post-prandial state [@Rhee2011a].
Because TGFA were only measured at the baseline visit, we cannot
investigate whether there are concomitant changes in TGFA and the metabolic
measures over time. However, to optimally use GEE to analyze the data and for
interpretation, we used the model to infer that a given value of TGFA could
predict values of insulin sensitivity or beta-cell function over a 6 year
period. This in our view is a strength of our analysis, as it reduces the chance
of reverse causality given the tight integration of the glucose and fatty acid
metabolism pathways, as well as maximizes the specific usage of the GEE
modeling.

PLS is a well-established technique for constructing predictive models of high
dimensionality data structures (i.e. fatty acid composition), however a 
limitation is that the initial models analyzed through PLS and the final 
computed scores are not able to control for potential confounders and other 
effect modifiers. PLS is also not able to handle longitudinal data so only 
the baseline visit was used in the PLS analysis, although we analyzed the extracted
scores using the GEE modeling to overcome this limitation and observed
concordant results between the PLS and GEE analyses.

Our study has several notable strengths, including the longitudinal design and
the use of advanced statistical techniques for data analysis. These statistical
techniques take advantage of the longitudinal data to allow appropriate
investigation of temporal relationships and are able to handle the
multidimensional nature of the data. Lastly, our cohort contains highly detailed
and comprehensive variable measurements for the fatty acids and outcomes,
of which were collected at each visit.

## Conclusion

In conclusion, we found that a TGFA composition containing higher proportions of
14:0, 14:1n-7, 16:0, and 16:1n-7 associated strongly with lower insulin
sensitivity and (more moderately) with lower beta-cell function. We also found
that most other TGFA (e.g. 20:0, most omega 6 and 9 TGFA, and 20:5n-3)
associated positively with insulin sensitivity. Only a few TGFA associated
positively and consistently with beta-cell function (e.g. 18:1n-7, 20:2n-6). 
These results provide more insight into how individual TGFA contribute to the
pathogenesis of T2D while reinforcing the importance and value of clinically
measured total TG as an indicator of metabolic health.

# Acknowledgements/grant support

The authors thank Jan Neuman, Paula Van Nostrand, Stella Kink, Nicole Rubio, and Annette
Barnie of the Leadership Sinai Centre for Diabetes, Mount Sinai Hospital,
Toronto, Canada and Sheila Porter and Mauricio Marin of the Centre for Studies
in Family Medicine, University of Western Ontario, London, Canada for their 
expert technical assistance and dedication in their work for PROMISE.

## Funding

This study was supported by grants from the Canadian Diabetes Association (CDA),
the Canadian Institutes for Health Research, and the University of
Toronto Banting and Best Diabetes Centre; LWJ was supported by a CDA
Doctoral Student Research Award; RR is supported by a Heart and Stroke
Foundation of Ontario Mid-Career Investigator Award; SBH holds the
CDA Chair in National Diabetes Management and the Ian McWhinney Chair of Family
Medicine Studies at the University of Western Ontario; RBP holds a
Tier II Canada Research Chair in Brain Lipid Metabolism; AJH holds a Tier II
Canada Research Chair in Diabetes Epidemiology.

## Conflicts of interest

The authors report no potential conflicts of interest relevant to this study.

## Contribution statement

The authors had the following responsibility: LWJ conducted research, analyzed
data, and wrote the paper; RR, BZ, ZL, and SBH designed research, conducted
research, and provided essential materials (infrastructure and clinical
resources); RR, BZ, SBH, RPB, and AG provided intellectual feedback on the
paper; RPB and ZL conducted research, provided essential reagents and materials; AJH
designed research, assisted with interpretation, and provided intellectual
feedback on all versions of the paper; LWJ and AJH had primary responsibility
for final content. All authors read and approved the final manuscript.

# Tables

```{r table_basic}
project_data %>% 
    table_basic(caption = tab_basic)
```

Note: Tables values as either median (interquartile range), mean (SD), or n (percent).

# Figures

```{r fig_tagfa, fig.cap=fig_tagfa}
project_data %>% 
    plot_tagfa()
```

```{r fig_heatmap, fig.cap=fig_heatmap}
plot_heatmap(cor_results$conc)
```

```{r fig_heatmap_mol, fig.cap=fig_heatmap_mol}
plot_heatmap(cor_results$pct, unit = 'mol%')
```

```{r fig_heatmap_tagfa, fig.cap=fig_heatmap_tagfa}
plot_heatmap(cor_results$tagfa, text = FALSE)
```

```{r fig_gee_unadj, fig.cap=fig_gee_unadj}
gee_results$unadj %>% 
    mutate(p.value = ifelse(p.value <= 0.05, 0.04, p.value)) %>% 
    plot_gee_main() +
    ggplot2::theme(legend.position = "none")
```

```{r fig_gee_adj, fig.cap=fig_gee_adj}
gee_results$adj %>% 
    mutate(p.value = ifelse(p.value <= 0.05, 0.04, p.value)) %>% 
    plot_gee_main() +
    ggplot2::theme(legend.position = "none")
```

```{r fig_pls, fig.cap=fig_pls, fig.height=8, fig.width=4}
p1 <- plot_pls_corrcomps(pls_results$homa2) + ggplot2::ggtitle('A: HOMA2-%S')
p2 <- plot_pls_corrcomps(pls_results$isi) + ggplot2::ggtitle('B: ISI')
p1 + p2 + plot_layout(ncol = 1)
```

# References